---
title: 'STA 601 Homework 9'
author: 'Lingyun Shao'
date: 'Nov. 21, 2018'
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
library(mvtnorm)
library(purrr)
library(MASS)
library(MCMCpack)
library(invgamma)
```

# 10.1
Reflecting random walks: It is often useful in MCMC to have a proposal distribution which is both symmetric and has support only on a certain region. For example, if we know $\theta$ > 0, we would like our proposal distribution $J(\theta_1|\theta_0)$ to have support on positive $\theta$ values. Consider the following proposal algorithm:

- sample $\tilde \theta\sim$ uniform($\theta_0-\delta, \theta_0 + \delta$);

- if $\tilde \theta < 0$, set $\theta_1 = -\tilde\theta$;

- if $\tilde \theta\geq0$, set $\theta_1 = \tilde\theta$.

In other words, $\theta_1$ = $|\tilde\theta|$. Show that the above algorithm draws samples from a symmetric proposal distribution which has support on positive values of $\theta$. It may be helpful to write out the associated proposal density $J(\theta_1|\theta_0)$ under the two conditions $\theta_0\leq\delta$ and $\theta_0>\delta$ separately.

\ 

Since we are getting our proposed sample $\theta_1=|\tilde \theta|\geq0$, it's straightforward that the proposal distribution has support on positive values.

\ 

For symmetry, notice that the symmetry in Metropolis algorithm is defined as $J(x|y)=J(y|x)$. According to the proposal algorithm given above, we can easily write down the proposal density as the following:

$$
J(\theta_1|\theta_0)=\begin{cases}\frac{1}{\delta},&\theta_0\leq\delta,0\leq\theta_1\leq\delta-\theta_0\\\frac{1}{2\delta},&\theta_0\leq\delta,\delta-\theta_0\leq\theta_1\leq\theta_0+\delta\\\frac{1}{2\delta},&\theta_0>\delta,\theta_0-\delta\leq\theta_1\leq\theta_0+\delta\\0,&o.w.\end{cases}
$$

For $\left\{\theta_0\leq\delta,0\leq\theta_1\leq\delta-\theta_0\right\}$, which is the same as $\left\{0<\theta_0,0<\theta_1,\theta_0+\theta_1\leq\delta\right\}$, it's the triangular area constrained by vertices $(0,0),(0,\delta),(\delta,0)$.

For $\left\{\theta_0\leq\delta,\delta-\theta_0\leq\theta_1\leq\theta_0+\delta\right\}$, it's the triangular area constrained by vertices $(0,\delta),(\delta,0),(0,2\delta)$.

For $\left\{\theta_0>\delta,\theta_0-\delta\leq\theta_1\leq\theta_0+\delta\right\}$, it's the area constrained by lines $\theta_0=\delta,\theta_1=\theta_0-\delta,\theta_1=\theta_0+\delta$ (the convex area).

If we regard the density as a bivariate function, and draw the areas for each condition respectively on a $\theta_1\ vs\ \theta_0$ coordinates, we can have the following density plot

```{r, fig.height=6.5}
plot(NA, xlim = c(-.5,2), ylim = c(-.5, 2), axes = FALSE,
     ylab = '', xlab = '')
abline(v=0)
abline(h=0)
segments(0,1,1,0)
polygon(x = c(0,0,1), 
        y = c(0,1,0),
        col = "steelblue")
segments(0,1,10,11)
segments(1,0,11,10)
polygon(x = c(0,1,12,11), 
        y = c(1,0,11,12),
        col = "slategray1")
text(0.25, 0.25, expression(frac(1,delta)))
text(1, 1, expression(frac(1,paste(2, delta))))
text(1, -0.15, expression(delta))
text(-0.15, 1, expression(delta))
text(2, -0.15, expression(theta[0]))
text(-0.15, 2, expression(theta[1]))
abline(a=0,b=1, lty = 2)
```

As is shown in the plot, we know that the support areas with probability densities are symmetric about $\theta_0=\theta_1$ line, which means that in the $\theta_1\ vs\ \theta_0$ space, the proposal distribution is completely identical if we switch $\theta_0$ and $\theta_1$. Therefore, we have verified that the proposal distribution $J(\theta_1|\theta_0)$ is symmetric, i.e. $J(\theta_1|\theta_0)=J(\theta_0|\theta_1)$.